### 2017-01-02 Chapters/experiments.tex
Finalize chap 3 suppinfo, finalize abstracts, remove eps files, version sent to Hidde and Hans

The DNA fragment containing the selection cassette was amplified using
two long primers annealing, respectively, to the region just before
the STOP codon of \textit{rpsB}, and to the beginning of \textit{gfp\_mut2}.
The selection cassette contains an antiobiotic resistance gene, kanamycin
(positive selection), and a gene coding for the CcdB toxin under the
control of the P\textsubscript{BAD} promoter (negative selection
in presence of arabinose). This cassette is referred below as \textit{kan-p\textsubscript{BAD}-ccdB}.

@@

\section{Supporting Information for Chapter~3}

@@

\subsubsection{Exhaustive list of the primers used...}

\paragraph{...for the \textit{gfpmut2} amplification.}

\begin{itemize}
\item[] \includegraphics[scale=1]{Fig/gfpmut2_L.png} \seqsplit{GTTCTCAGGATCTGGCTTCCCAGGCGGAAGAAAGCTTCGTAGAAGCTGAGCAGGAAAGGCGACAGGAGAGTAAAGGAGAAGAACTTTTCACTG}
(Length: 93)
\item[] \includegraphics[scale=1]{Fig/gfpmut2_R.png} \seqsplit{TGATGTTCTGGGGAATATAATTATTTGTATAGTTCATCCATGCC} (Length: 44)
\end{itemize}

\textbf{Rem.} The 50 first bases of the forward primer are supposed to anneal with the end of the \textit{rpsB} gene.
The 20 last bases of the reverse primer are supposed to anneal with the ccdB-end of the \textit{kan-P\textsubscript{BAD}-ccdB} cassette.

@@

\begin{itemize}
\item[] \includegraphics[scale=1]{Fig/mCherry_L.png} \seqsplit{GTTCTCAGGATCTGGCTTCCCAGGCGGAAGAAAGCTTCGTAGAAGCTGAGCAGGAAAGGCGACAGGAGACTAGCAAAAGATCCAAGGG}
(Length: 88)
\item[] \includegraphics[scale=1]{Fig/mCherry_R.png} \seqsplit{TGATGTTCTGGGGAATATAATTATTTGTACAGCTCATCCATG} (Length: 42)
\end{itemize}

\textbf{Rem.} The 50 first bases of the forward primer are supposed to anneal with the end of the \textit{rpsB} gene.
The 20 last bases of the reverse primer are supposed to anneal with the ccdB-end of the \textit{kan-P\textsubscript{BAD}-ccdB} cassette.

@@

\begin{itemize}
\item[] \includegraphics[scale=1]{Fig/cassette_L} \seqsplit{TGGATGAACTATACAAATAATTATATTCCCCAGAACATCAGG}
(Length: 42, used to assemble with gfp)
\item[] \includegraphics[scale=1]{Fig/cassette_L} \seqsplit{TGGATGAGCTGTACAAATAATTATATTCCCCAGAACATCAG}
(Length: 41, used to assemble with mCherry)
\item[] \includegraphics[scale=1]{Fig/cassette_R} \seqsplit{GAGCTTGCCGCCTTTCTGCAACTCGAACTATTTTGGGGGAGTTATCAAGCTTAGAAGAACTCGTCAAGAAGG}
(Length: 72, used for both)
\end{itemize}

\textbf{Rem.} The 20 first bases of the first forward primer are supposed to anneal with the end of the \textit{gfpmut2} gene.
The 20 first bases of the second forward primer are supposed to anneal with the end of the \textit{mCherry} gene.
The 50 last bases of the reverse primer are supposed to anneal with the region after \textit{rpsB} on the \textit{E.~coli} BW25113 chromosome.

@@

\begin{itemize}
\item[\textbf{\textit{rpsB-gfpmut2}}] \seqsplit{TTGTAACAGCTGCTGGGATTACACATGGCATGGATGAACTATACAAATAAGCTTGATAACTCCCCCAAAATAGTTCGAGTTGCAGAAAGGCGGCAAGCTC}
(Length: 100)
\item[\textbf{\textit{rpsB-mCherry}}] \seqsplit{GCGCGGAGGGTCGTCATTCTACCGGTGGCATGGATGAGCTGTACAAATAAGCTTGATAACTCCCCCAAAATAGTTCGAGTTGCAGAAAGGCGGCAAGCTC}
(Length: 100)
\end{itemize}

@@

The following pages contain the full sequences of the \textit{gfpmut2-cassette} and \textit{mCherry-cassette} constructions, inserted on the \textit{E.~coli} BW25113 chromosome at the region after the \textit{rpsB} gene (see Fig.~\ref{fig:Cloning}).
The primers annealing positions are indicated using the same color code as in the description of the primers above.
The primers reported in the original publication of Bakshi \textit{et al.}~\cite{bakshi_superresolution_2012} are indicated for information.

\includepdf[pages={1-7}]{./Fig/gfp-mut2-cassette-after-rpsB-gene-sequence.pdf}
\includepdf[pages={1-6}]{./Fig/mcherry-cassette-after-rpsB-gene-sequence.pdf}

@@

The strains were constructed using the protocol described in Material and Methods~\ref{sec:methods_strain}.
The downstream region of the \textit{rpsB} gene was amplified, and sent for sequencing.
Alignments of the results showed they exhibit the expected sequences, except for a couple of single-base mutations that do not impede the protein fluorescence.
The raw results of the sequencing are reproduced below.

@@

The \textit{rpsB-gfp} and \textit{rpsB-mCherry} strains were characterized on different media using a Tecan microplate reader.
Strains were grown in M9 minimal medium at 37$^\circ$C in 96-well microplates.
The absorbance and fluorescence were measured approximately every minute for up to 24~h.
In a single experiment, this allowed to construct up to 96 growth curves like presented in Fig.~\ref{fig:growth_curve}.

In Fig.~\ref{fig:suppinfo_caract}, we can see the results obtained on M9 supplemented with 0.2\% glucose.
As for the construction of Bakshi \textit{et al.}~\cite{bakshi_superresolution_2012}, our strains possess a wild-type growth rate (Fig.~\ref{fig:suppinfo_caract}A), indicating that the gfp-tagging or mCherry-tagging of the ribosomal S2 subunit does not impede the functioning of the ribosome.
It is known that \textit{E.~coli} cultures exhibit a high autofluorescence in the GFP bandwidth~\cite{mihalcescu_green_2015}, but the fluorescence of the \textit{rpsB-gfp} strain is at least 4-time higher than the \textit{wild-type} (Fig.~\ref{fig:suppinfo_caract}B).
Contrary to green autofluorescence, red autofluorescence of \textit{E.~coli} cultures is very low (Fig.~\ref{fig:suppinfo_caract}B) so the signal level for mCherry was even less a problem.
Overall, both strains possess enough fluorescence to allow quantification.

In Fig.~\ref{fig:suppinfo_caract}C, we display the ratio between the (corrected) fluorescence of the strain and the absorbance.
As a first approximation, this ratio can be used as a proxy for the fluorescence concentration in the cells, therefore for the ribosome concentration.
The instability at the beginning of the experiment (before 400~min) is characteristic of the difficulty to obtain robust estimations when the absorbance of the culture is low~\cite{zulkower_robust_2015}.
The RFU/Abs ratio is stable in the interval [400,600] for both \textit{rpsB-gfp} and \textit{rpsB-mCherry} strains, and can be interpreted as exponential steady-state growth.
Note that this short time interval (a couple of generations) of readability before approaching the stationary phase is one of the main problem preventing the use of batch conditions to obtain robust data about growth transitions.

@@

\begin{figure}[p]
\centering
\includegraphics[scale=1]{./Fig/suppinfo_caract}
\caption{
\textbf{Growth curves in M9~0.2\% glucose for the \textit{rpsB-gfp} and \textit{rpsB-mCherry} strains.}
Growth and measurements were performed in a 96-well microplate placed in a Tecan infinite 200 pro reader thermostated at 37$^\circ$C.
We monitored 4 wells inoculated with the WT strain, the \textit{rpsB-gfp} strain, and the \textit{rpsB-mCherry} strain, for a total of 12 wells each containing 150~$\mu$L of growth medium.
Curves were shifted in time to correct for pipetting variability in the inoculation process (-50 min for WT, and -20 min for \textit{rpsB-mCherry}).
\textit{(A)}~The light absorbance as measured by the Tecan microplate reader, corrected for background by removing well containing clean M9.
The superposition of the curves indicates similar growth rates between the WT and the modified strains.
\textit{(B)}~The fluorescence measured in the wells, expressed in Relative Fluorescence Units (RFU).
Green fluorescence (485nm excitation, 535nm reception) is measured for the WT (dashed lines) and the \textit{rpsB-gfp} (solid lines) strains.
Red fluorescence (560nm excitation, 635nm reception) is measured for the WT (dashed lines) and the \textit{rpsB-mCherry} (solid lines) strains.
As can be seen, fluorescence levels of the modified strains (solid lines) are far above the autofluorescence measured on the WT strain.
\textit{(C)}~The ratio of fluorescence over absorbance, a proxy for the fluorescence concentration in the cells.
Autofluorescence background was corrected by removing the autofluorescence measured on the WT wells.
We observe a strange increase in fluorescence concentration for the \textit{rpsB-mCherry} strain at stationary phase, something that is not visible on the \textit{rpsB-gfp} strain.
Before each measurement cycle, the following procedure was applied: shaking (Orbital 6mm) for 30~s, shaking (Linear 6mm) for 30~s, waiting for 5~s.
}
\label{fig:suppinfo_caract}
\end{figure}

In order to explain these strange dynamics, we set-up an experiment to evaluate the degradation and maturation rates of GFP and mCherry.
The procedure was to abruptly stop the protein synthesis rate in mid-exponential phase, and observe the evolution of the fluorescence in a condition with no synthesis of new reporter proteins.
The estimation was performed using the following model:
\begin{eqnarray}
\frac{dP_m}{dt} &=& K \cdot P - (\mu+D) \cdot P_m, \\
\frac{dP}{dt} &=& f(t) - (\mu+D+K) P,
\end{eqnarray}
where $P_m$ is the concentration of mature (fluorescent) proteins, $P$ is the concentration of immature (non-fluorescent) proteins, $K$ is the maturation rate, $D$ is the degradation rate (assumed identical for mature and immature proteins), $f(t)$ the protein synthesis rate, and $\mu$ the growth rate.
By stopping all protein synthesis in the bacteria, we obtain $f(t) = \mu = 0$, hence the model becomes:
\begin{eqnarray}
\frac{dP_m}{dt} &=& K \cdot P - D \cdot P_m, \\
\frac{dP}{dt} &=& - (D+K) P.
\end{eqnarray}
This system of differential equations can be analytically solved into:
\[
P_m(t) = \left(P_m(0) + P(0) \left( 1 - e^{-Kt} \right) \right) e^{-Dt}
\]
with $P_m(0)$ and $P(0)$ the initial concentration of mature proteins and immature proteins, respectively.
By normalizing by $P_m(0)$ and taking the log of both sides, we can rewrite it as
\begin{equation}
\log \frac{P_m(t)}{P_m(0)} = - Dt + \log \left(1 + \frac{P(0)}{P_m(0)} \left( 1 - e^{-Kt} \right) \right).
\label{eq:matur_deg}
\end{equation}
Since growth is stopped, the bacterial volume is constant and then $\frac{P_m(t)}{P_m(0)}$ can be easily measured by directly taking the fluorescence normalized by its initial value, while the maturation rate $K$, the degradation rate $D$, and the initial fraction of immature proteins $\frac{P(0)}{P_m(0)}$ are free parameters that can be fitted on fluorescence time series.
The curves used for the estimation are represented in Fig.~\ref{fig:suppinfo_matur} and the results are reported in Tab.~\ref{tab:suppinfo_matur}.
As expected, both reporter proteins are stable and exhibit a long half-life in exponential phase (>24~h).
The maturation rate of GFP was too fast to obtain numerical value.
However, the maturation rate of mCherry is rather slow, with a half-maturation time in the order of 30~min (time needed to mature 50\% of a pool of mCherry).
Even though degradation and maturation rate can be slightly different in stationary phase, the estimated value is not sufficient to explain the increase in fluorescence observed in Fig.~\ref{fig:suppinfo_caract}.

\begin{figure}[tb]
\centering
\includegraphics[scale=1]{./Fig/suppinfo_matur}
\caption{
\textbf{Estimation of the maturation and degradation rates of the reporter proteins in the \textit{rpsB-gfp} and \textit{rpsB-mCherry} strains.}
The strains were grown in the same condition as in Fig.~\ref{fig:suppinfo_caract}, except a massive dose of Chloramphenicol was added at t=0 (vertical dashed grey line).
The fluorescence was normalized by the value at t=0 and the degradation and maturation rates were estimated by using the model in Eq.~\ref{eq:matur_deg}.
Parameters of the best fit (solid black line) are reported in Tab.~\ref{tab:suppinfo_matur}.
Data are the result of the aggregation of 9 and 8 independent growth curves for \textit{rpsB-gfp} and \textit{rpsB-mCherry}, respectively.
}
\label{fig:suppinfo_matur}
\end{figure}

@@

We estimated the noise in the RFU/pixel/cell by using the data points just before upshift (Fig.~\ref{fig:noise_poi}).
Since they have been growing for 20~h on acetate in the microfluidic device, bacteria are assumed in balanced growth in this region of $\sim$2~hours (roughly 23-24 points depending on the well of interest).
For this reason, ribosome concentration, and so fluorescence concentration, are expected constant.
Fig.~\ref{fig:noise_poi} represents the distribution of points in this region for a particular cell, that can be approximated by a Gaussian distribution.

\begin{figure}[p]
\centering
\includegraphics[scale=1]{./Fig/noise_poi}
\caption{
\textbf{Point of interest for the noise estimation.}
\textit{(Top graph)} RFU/pixel/cell data points for the cell at the bottom of the well labeled X3Y2,W2.
Noise estimation was performed on the points just before the upshift (red points), were the bacteria is assumed to grow at steady state.
\textit{(Bottom graph)} Cumulative density function (CDF, in red) and probability density function (CDF, in gray) of the points highlighted in red on the top graph.
They are visually compared with CDF (solid green line) and PDF (dashed green line) of a Gaussian fit with mean 548.88 and standard deviation 17.316.
}
\label{fig:noise_poi}
\end{figure}

@@

Noise characteristics were computed by normalizing and aggregating data for the 45 normal cells presented in Fig.~\ref{fig:noise_distrib_mean_std}.
Two scenarii were considered.
First, the noise model can be considered additive, which means we have the measurement model presented in Section~\ref{sec:estim_gr_ra}:
\begin{equation}
F(t_k) = \gamma r (t_k) + \eta_k,\\
\end{equation}
were $F(t_k)$ is the fluorescence measured at the time step $t_k$, $r (t_k)$ is the true ribosomal concentration at this time step, $\gamma$ is an unknown factor, and $\eta_k$ is the measurement noise.
In this scenario, for a time independent ribosome concentration $r$ (steady state), the ribosome concentration is constant and equal to its mean on the interval of interest (noted $m_k(\cdot)$).
Given that
\begin{equation}
m_k (F) = m_k (\gamma r) + m_k(\eta_k) = \gamma m_k (r),\\
\end{equation}
the noise residues are given by
\begin{equation}
\eta_k = F(t_k) - m_k (F). \label{eq:noise_add}
\end{equation}
In other words, we can normalize the data points of each cell by removing their mean in order to aggregate the noise residues and get an estimation of the noise level (Fig.~\ref{fig:noise_estim_add_mul}, top panel).

Secondly, the noise model can be considered multiplicative.
The noise model is than different than presented in Section~\ref{sec:estim_gr_ra}, and can be re-written:
\begin{equation}
F(t_k) = \gamma r (t_k) \cdot (1 + \lambda_k),
\end{equation}
were $\gamma r (t_k) \cdot \lambda_k$ is the measurement noise, proportional to the value measured.
In that scenario, with the same notations as above, the noise $\lambda_k$ is given by
\begin{equation}
\lambda_k = \frac{F(t_k)}{m_k (F)} - 1, \label{eq:noise_mul}
\end{equation}
which means we can normalize data points from different cells by dividing by the mean and removing 1 (Fig.~\ref{fig:noise_estim_add_mul}, bottom panel).

Both models were considered and are represented in Fig.~\ref{fig:noise_estim_add_mul}.
Despite the correlation identified in Fig.~\ref{fig:noise_distrib_mean_std}, the means of the residues in both models are very close to zero, and the distributions are well approximated by a Gaussian distribution.
For this reason, we decided in the main analysis to stick with an additive noise, the main reason being that the implementation used for the Kalman smoothing procedure are faster and more stable with an additive noise model.
We thus assumed an additive white Gaussian noise with mean 0 and standard deviation 28.29 for the simulation of the synthetic data and the signal reconstruction through the Kalman smoothing procedure.
As discussed in Section~\ref{sec:chap3_discussion}, the only way to decisively choose a noise model over the other would have been to obtain two different steady-state growths for the each cell.
Unfortunately, the presented time series were not long enough to obtain a new steady state on glucose, something that will be corrected in future instances of the experiment.

@@

To help the reader get a better idea on the analysis, we display below the complete reconstructions for 5 cells.
In Figs~\ref{fig:cell1}-\ref{fig:cell5}, the image on the left is the last image analyzed for the corresponding well.
The bacteria either at the top or the bottom (depending on the orientation in the device) was manually segmented by selecting two pixels at the poles on the fluorescence images (red cross).
A 6-pixel-wide rectangular mask was computed for each image, resulting in the masked image on the right.
In each of the graphs, the black and green points represents data points, while the red solid lines are the results of the signal reconstruction \textit{via} the Kalman smoothing procedure, the parameters of which are described in Material and Methods~\ref{sec:meth_kalman} and Figs~\ref{fig:growth_rate_estimation} and \ref{fig:gene_activity}.
Specific comments on each cell are given in the caption of Figs~\ref{fig:cell1}-\ref{fig:cell5}.

>>>>

The DNA fragment containing the selection cassette was amplified using
two long primers annealing, respectively, downstream of \textit{rpsB} (starting just after
the STOP codon), and to the end of \textit{gfp\_mut2}.
The selection cassette contains an antibiotic resistance gene, kanamycin
(positive selection), and a gene coding for the CcdB toxin under the
control of the \textit{p\textsubscript{BAD}} promoter (negative selection
in presence of arabinose). This cassette is referred below as \textit{kan-p\textsubscript{BAD}-ccdB}.

@@

\clearpage
\section{Supporting Information for Chapter~3}

@@

\subsubsection{Exhaustive list of the primers used}

\paragraph{...for the \textit{gfp\_mut2} amplification.}

\begin{itemize}
\item[] \includegraphics[scale=1]{Fig/gfpmut2_L.png} \seqsplit{GTTCTCAGGATCTGGCTTCCCAGGCGGAAGAAAGCTTCGTAGAAGCTGAGCAGGAAAGGCGACAGGAGAGTAAAGGAGAAGAACTTTTCACTG}
(Length: 93)\\
\textit{The 50 first bp anneals with the C-terminus of \textit{rpsB} (just before the STOP codon).
The 18 next bp code for the linker (see Material and Methods~\ref{sec:methods_strain}).}
\item[] \includegraphics[scale=1]{Fig/gfpmut2_R.png} \seqsplit{TGATGTTCTGGGGAATATAATTATTTGTATAGTTCATCCATGCC} (Length: 44)\\
\textit{The 20 last bp anneals with the end of the \textit{kan-p\textsubscript{BAD}-ccdB} cassette.}
\end{itemize}

@@

\begin{itemize}
\item[] \includegraphics[scale=1]{Fig/mCherry_L.png} \seqsplit{GTTCTCAGGATCTGGCTTCCCAGGCGGAAGAAAGCTTCGTAGAAGCTGAGCAGGAAAGGCGACAGGAGACTAGCAAAAGATCCAAGGG}
(Length: 88)
\item[] \includegraphics[scale=1]{Fig/mCherry_R.png} \seqsplit{TGATGTTCTGGGGAATATAATTATTTGTACAGCTCATCCATG} (Length: 42)\\
\textit{Same design as for gfp\_mut2, except it amplifies mCherry.}
\end{itemize}

@@

\begin{itemize}
\item[] \includegraphics[scale=1]{Fig/cassette_L} \seqsplit{TGGATGAACTATACAAATAATTATATTCCCCAGAACATCAGG}
(Length: 42, used to assemble with gfp)\\
\textit{The 20 first bp anneals with the end of gfp\_mut2.}
\item[] \includegraphics[scale=1]{Fig/cassette_L} \seqsplit{TGGATGAGCTGTACAAATAATTATATTCCCCAGAACATCAG}
(Length: 41, used to assemble with mCherry)\\
\textit{The 20 first bp anneals with the end of mCherry.}
\item[] \includegraphics[scale=1]{Fig/cassette_R} \seqsplit{GAGCTTGCCGCCTTTCTGCAACTCGAACTATTTTGGGGGAGTTATCAAGCTTAGAAGAACTCGTCAAGAAGG}
(Length: 72, used for both)\\
\textit{The 50 last bp anneals downstream rpsB (just before the STOP codon).}
\end{itemize}

@@

\begin{itemize}
\item[\textbf{\textit{rpsB-gfp}}] \seqsplit{TTGTAACAGCTGCTGGGATTACACATGGCATGGATGAACTATACAAATAAGCTTGATAACTCCCCCAAAATAGTTCGAGTTGCAGAAAGGCGGCAAGCTC}
(Length: 100)
\item[\textbf{\textit{rpsB-mCherry}}] \seqsplit{GCGCGGAGGGTCGTCATTCTACCGGTGGCATGGATGAGCTGTACAAATAAGCTTGATAACTCCCCCAAAATAGTTCGAGTTGCAGAAAGGCGGCAAGCTC}
(Length: 100)
\end{itemize}

@@

The following pages contain the full annotated sequences of the region downstream \textit{rpsB} after the insertion of \textit{gfp\_mut2-cassette} and \textit{mCherry-cassette}, respectively (third line in Fig.~\ref{fig:Cloning}).
The primers annealing positions are indicated using the same color code as in the description of the primers above.
The primers reported in the publication of Bakshi \textit{et al.}~\cite{bakshi_superresolution_2012} are indicated for information.

\includepdf[pages={1-7}]{./Fig/gfp_mut2-cassette-after-rpsB-E-coli-sequence.pdf}
\includepdf[pages={1-7}]{./Fig/mCherry-cassette-after-rpsB-E-coli-sequence.pdf}

@@

After the strains were made using the protocol described in Material and Methods~\ref{sec:methods_strain}, the downstream region of the \textit{rpsB} gene was amplified and sent for sequencing.
Alignments of the results showed they exhibit the expected sequences, except for a couple of single-base mutations.
The raw results of the sequencing are reproduced below.

@@

The \textit{rpsB-gfp} and \textit{rpsB-mCherry} strains were characterized on different media using a Tecan microplate reader.
Strains were grown in M9 minimal medium at 37$^\circ$C in 96-well microplates.
The absorbance and fluorescence were measured approximately every minute for up to 24~h.
In a single experiment, this allowed to generate up to 96 growth curves like the one schematized in Fig.~\ref{fig:growth_curve}.

In Fig.~\ref{fig:suppinfo_caract}, we show growth curves obtained on M9 supplemented with 0.2\% glucose.
As for the construction of Bakshi \textit{et al.}~\cite{bakshi_superresolution_2012}, our strains possess a wild-type growth rate (Fig.~\ref{fig:suppinfo_caract}A), indicating that the gfp-tagging or mCherry-tagging of the ribosomal S2 subunit does not impede the functioning of the ribosome.
It is known that \textit{E.~coli} cultures exhibit a high autofluorescence in the GFP bandwidth~\cite{mihalcescu_green_2015}, but the fluorescence of the \textit{rpsB-gfp} strain is at least 4-time higher than the \textit{wild-type} (Fig.~\ref{fig:suppinfo_caract}B).
Contrary to green autofluorescence, red autofluorescence of \textit{E.~coli} cultures is very low (Fig.~\ref{fig:suppinfo_caract}B) so the corrected signal level for mCherry is even more stronger.
Overall, both strains possess enough fluorescence to allow quantification at the population level, all the more so at the cell level because most part of the autofluorescence is concentrated in the medium outside the cell~\cite{mihalcescu_green_2015}.

In Fig.~\ref{fig:suppinfo_caract}C, we display the ratio between the (corrected) fluorescence of the strain and the absorbance (corrected means that we substracted the autofluorescence of the wild-type strain).
As a first approximation, this ratio can be used as a proxy for the fluorescence concentration in the cells, therefore for the ribosome concentration.
The instability at the beginning of the experiment (before 400~min) is characteristic of the difficulty to obtain robust estimations when the absorbance of the culture is low~\cite{zulkower_robust_2015}.
The RFU/Abs ratio is stable in the interval [400,600]~min for both \textit{rpsB-gfp} and \textit{rpsB-mCherry} strains, and can be interpreted as exponential steady-state growth.
Note that this short time interval (a couple of generations) of readability before approaching the stationary phase is one of the main problem preventing the use of batch conditions to obtain robust data about growth transitions.
Indeed, most of our attempt to perform robust growth transitions in this region were unsuccessful.

@@

\begin{figure}[p]
\centering
\includegraphics[scale=1]{./Fig/suppinfo_caract}
\caption{
\textbf{Growth curves in M9~0.2\% glucose for the \textit{rpsB-gfp} and \textit{rpsB-mCherry} strains.}
Growth and measurements were performed in a 96-well microplate placed in a Tecan infinite 200 pro reader thermostated at 37$^\circ$C.
We monitored 4 wells inoculated with the WT strain, the \textit{rpsB-gfp} strain, and the \textit{rpsB-mCherry} strain, for a total of 12 wells each containing 150~$\mu$L of growth medium.
Curves were shifted in time to correct for pipetting variability in the inoculation process (-50 min for WT, and -20 min for \textit{rpsB-mCherry}).
\textit{(A)}~The light absorbance as measured by the Tecan microplate reader, corrected for background by removing M9 absorbance.
The superposition of the curves indicates similar growth rates between the WT and the modified strains.
\textit{(B)}~The fluorescence measured in the wells, expressed in Relative Fluorescence Units (RFU).
Green fluorescence (485~nm excitation, 535~nm emission) is measured for the WT (dashed lines) and the \textit{rpsB-gfp} (solid lines) strains.
Red fluorescence (560~nm excitation, 635~nm emission) is measured for the WT (dashed lines) and the \textit{rpsB-mCherry} (solid lines) strains.
As can be seen, fluorescence levels of the modified strains are far above the autofluorescence measured on the WT strain.
\textit{(C)}~The ratio of fluorescence over absorbance, a proxy for the fluorescence concentration in the cells.
Autofluorescence background was corrected by removing the autofluorescence measured on the WT wells.
We observe a strange increase in fluorescence concentration for the \textit{rpsB-mCherry} strain at stationary phase, something does not occur on the \textit{rpsB-gfp} strain.
Before each measurement cycle, the following procedure was applied: shaking (Orbital 6mm) for 30~s, shaking (Linear 6mm) for 30~s, waiting for 5~s.
}
\label{fig:suppinfo_caract}
\end{figure}

In order to explain these strange dynamics, we set-up an experiment to evaluate the degradation and maturation rates of GFP and mCherry in our strains.
The procedure was to instantly stop all protein synthesis in mid-exponential phase using an antibiotic, then to observe the evolution of the fluorescence in a condition with no synthesis of new reporter proteins.
The estimation was performed using the following model:
\begin{eqnarray}
\frac{dP_m}{dt} &=& K \cdot P - (\mu+D) \cdot P_m, \\
\frac{dP}{dt} &=& f(t) - (\mu+D+K) P,
\end{eqnarray}
where $P_m$ is the concentration of mature (fluorescent) proteins, $P$ is the concentration of immature (non-fluorescent) proteins, $K$ is the maturation rate, $D$ is the degradation rate (assumed identical for mature and immature proteins), $f(t)$ the protein synthesis rate, and $\mu$ the growth rate.
By stopping all protein synthesis in the bacteria, we obtain $f(t) = \mu = 0$, hence the model becomes:
\begin{eqnarray}
\frac{dP_m}{dt} &=& K \cdot P - D \cdot P_m, \\
\frac{dP}{dt} &=& - (D+K) P.
\end{eqnarray}
This system of differential equations can be analytically solved into:
\[
P_m(t) = \left(P_m(0) + P(0) \left( 1 - e^{-Kt} \right) \right) e^{-Dt}
\]
with $P_m(0)$ and $P(0)$ the initial concentrations of mature proteins and immature proteins, respectively.
By dividing by $P_m(0)$ and taking the log of both sides, we can rewrite it as
\begin{equation}
\log \frac{P_m(t)}{P_m(0)} = - Dt + \log \left(1 + \frac{P(0)}{P_m(0)} \left( 1 - e^{-Kt} \right) \right).
\label{eq:matur_deg}
\end{equation}
Since growth is stopped, the bacterial volume is constant and then $\frac{P_m(t)}{P_m(0)}$ can be easily measured by directly taking the fluorescence normalized by its initial value, while the maturation rate $K$, the degradation rate $D$, and the initial fraction of immature proteins $\frac{P(0)}{P_m(0)}$ are free parameters that can be fitted on fluorescence time series.
The curves used for the estimation are represented in Fig.~\ref{fig:suppinfo_matur} and the results are reported in Tab.~\ref{tab:suppinfo_matur}.
As expected, both reporter proteins are stable and exhibit a long half-life in exponential phase (>24~h).
The maturation rate of GFP was too fast to obtain numerical value.
However, the maturation rate of mCherry is rather slow, with a half-maturation time in the order of 30~min (time needed to mature 50\% of a pool of mCherry).
Even though degradation and maturation rate can be slightly different in stationary phase, the estimated value are not sufficient to explain the rapid increase in fluorescence observed in Fig.~\ref{fig:suppinfo_caract}.

\begin{figure}[tb]
\centering
\includegraphics[scale=1]{./Fig/suppinfo_matur}
\caption{
\textbf{Estimation of the maturation and degradation rates of the reporter proteins in the \textit{rpsB-gfp} and \textit{rpsB-mCherry} strains.}
The strains were grown in the same condition as in Fig.~\ref{fig:suppinfo_caract}, except that a massive dose of Chloramphenicol was added in mid-exponential phase ($t=0$, vertical dashed grey line).
The fluorescence was normalized by the value at $t=0$ and the degradation and maturation rates were estimated by using the model in Eq.~\ref{eq:matur_deg}.
Parameters of the best fit (solid black line) are reported in Tab.~\ref{tab:suppinfo_matur}.
Data are the result of the aggregation of 9 and 8 independent growth curves for \textit{rpsB-gfp} and \textit{rpsB-mCherry}, respectively.
}
\label{fig:suppinfo_matur}
\end{figure}

@@

We estimated the noise in the RFU/pixel/cell measurements by using the data points just before the upshift (Fig.~\ref{fig:noise_poi}).
Since they have been growing for 20~h on acetate in the microfluidic device, bacteria are assumed in balanced growth in this region of $\sim$2~hours (roughly 23-24 points depending on the well of interest).
For this reason, ribosome concentration, and so fluorescence concentration, are expected constant.
Fig.~\ref{fig:noise_poi} shows the distribution of points in this region, that can be approximated by a Gaussian distribution, for a particular cell.

\begin{figure}[p]
\centering
\includegraphics[scale=1]{./Fig/noise_poi}
\caption{
\textbf{Points of interest for the noise estimation.}\newline
\textit{(Top graph)} RFU/pixel/cell data points for the cell at the bottom of the well labeled X3Y2,W2.
Noise estimation was performed on the points just before the upshift (red points), were the bacteria is assumed to grow at steady state.
\textit{(Bottom graph)} Cumulative density function (CDF, in red) and probability density function (PDF, in gray) of the points highlighted in red on the top graph.
They are visually compared with CDF (solid green line) and PDF (dashed green line) of a Gaussian fit with mean 548.88 and standard deviation 17.316.
}
\label{fig:noise_poi}
\end{figure}

@@

Noise characteristics were computed by normalizing and aggregating data for the 45 normal cells presented in Fig.~\ref{fig:noise_distrib_mean_std}.
Two scenarii were considered.
First, the noise model can be considered additive, which means we have the measurement model presented in Section~\ref{sec:estim_gr_ra}:
\begin{equation}
F(t_k) = \gamma r (t_k) + \eta_k,\\
\end{equation}
were $F(t_k)$ is the fluorescence measured at the time step $t_k$, $r (t_k)$ is the true ribosomal concentration at this time step, $\gamma$ is an unknown factor, and $\eta_k$ is the measurement noise.
In this scenario, for a time independent ribosome concentration $r$ (\textit{i.e.} at steady state), the ribosome concentration is equal to its mean on the interval of interest (noted $m_k(\cdot)$).
Given that
\begin{equation}
m_k (F) = m_k (\gamma r) + m_k(\eta_k) = \gamma m_k (r),\\
\end{equation}
the noise residues are given by
\begin{equation}
\eta_k = F(t_k) - m_k (F). \label{eq:noise_add}
\end{equation}
In other words, we can normalize the data points of each cell by removing their mean in order to aggregate the noise residues and get an estimation of the noise level (Fig.~\ref{fig:noise_estim_add_mul}, top panel).

Alternatively, the noise model can be considered multiplicative.
The noise model is than different than presented in Section~\ref{sec:estim_gr_ra}, and can be re-written:
\begin{equation}
F(t_k) = \gamma r (t_k) \cdot (1 + \lambda_k),
\end{equation}
were $\gamma r (t_k) \cdot \lambda_k$ is the measurement noise, proportional to the value measured.
In that scenario, with the same notations as above, the noise $\lambda_k$ is given by
\begin{equation}
\lambda_k = \frac{F(t_k)}{m_k (F)} - 1, \label{eq:noise_mul}
\end{equation}
which means we can normalize data points from different cells by dividing by the mean and removing 1 (Fig.~\ref{fig:noise_estim_add_mul}, bottom panel).

Both models were considered and are represented in Fig.~\ref{fig:noise_estim_add_mul}.
Despite the correlation identified in Fig.~\ref{fig:noise_distrib_mean_std}, the means of the residues in both models are very close to zero, and the distributions are well approximated by a Gaussian distribution.
For this reason, we decided in the main analysis to stick with an additive noise, the main reason being that the implementation used for the Kalman smoothing procedure are faster and more stable with an additive noise model.
We assumed an additive white Gaussian noise with mean 0 and standard deviation 28.29 for the simulation of the synthetic data and the signal reconstruction through the Kalman smoothing procedure.
As discussed in Section~\ref{sec:chap3_discussion}, the only way to decisively choose a noise model over the other would have been to obtain two different steady-state growths for each cell.
Unfortunately, the presented time series were not long enough to obtain a new steady state on glucose, something that will be corrected in future instances of the experiment.

@@

In order to get a more visual idea of the results of the analysis, we display below the complete signal reconstructions for 5 cells.
In Figs~\ref{fig:cell1}-\ref{fig:cell5}, the image on the left is the last image analyzed for the corresponding well.
The bacteria either at the top or the bottom (depending on the orientation in the device) was manually segmented by selecting two pixels at the poles on the fluorescence images (red cross).
A 6-pixel-wide rectangular mask was computed for each image, resulting in the masked image on the right.
In each of the graphs, the black and green points represent data points, while the red solid lines are the results of the signal reconstruction \textit{via} the Kalman smoothing procedure.
The parameters of the Kalman smoothing are described in Material and Methods~\ref{sec:meth_kalman} and Figs~\ref{fig:growth_rate_estimation} and \ref{fig:gene_activity}.
Specific comments on each cell are given in the caption of Figs~\ref{fig:cell1}-\ref{fig:cell5}.

\begin{figure}[p]
\centering
\includegraphics[scale=1]{./Fig/cell1}
\caption{
\textbf{Complete analysis for the cell located at the bottom of the well labelled "X1Y1, W11".}
(Figure description available in the introduction of \ref{S8_Text}.)\newline
Growth rate and resource allocation are particularly unstable at the beginning of the experiment, but seems to have stabilized before the upshift.
Oscillations in the RFU/pixel/cell signal are clearly visible, and result in oscillations in the resource allocation signal reconstruction.
}
\label{fig:cell1}
\end{figure}
